Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-19 paper/icml_drau/dauc.tex

%! TEX root = icml_drau.tex
\begin{algorithmic}
	\STATE \textbf{Inputs:}
	\STATE $\theta$ and $\psi$, parameters of
	$V_{\theta}$ and $\bar{A}_{\psi}$.
	\STATE $\pi^{\text{explore}}$ and $\nu_\deltat$ defining an exploration policy.
	\STATE \textbf{opt}$_1$, \textbf{opt}$_2$, $\alpha_1 \deltat$ and $\alpha_2 \deltat$ optimizers and learning rates.
	\STATE $\mathcal{D}$, buffer of transitions $(s, a, r, d, s')$, with $d$ the episode termination signal.
	\STATE $\deltat$ and $\gamma$, time discretization and discount factor.
	\STATE \textbf{nb\_epochs} number of epochs.
	\STATE \textbf{nb\_steps}, number of steps per epoch.
	\STATE
	\STATE Observe initial state $s^0$
	\STATE $t \gets 0$
	\FOR {$e=0, \textbf{nb\_epochs}$}
	\FOR {$j=1, \textbf{nb\_steps}$}
	\STATE $a^k \leftarrow \pi^{\text{explore}}(s^k, \nu^k_\deltat)$.
	\STATE Perform $a^k$ and observe $(r^{k+1}, d^{k+1}, s^{k+1})$.
	\STATE Store $(s^k, a^k, r^{k+1}, d^{k+1}, s^{k+1})$ in $\mathcal{D}$.
	\STATE $k \gets k + 1$
	\ENDFOR
	\FOR {$k=0, \text{nb\_learn}$}
	\STATE \text{Sample a batch of $N$ random transitions from $\mathcal{D}$}
	\STATE $Q^i \gets V_{\theta}(s^i) + \deltat\hspace{-.17em}\left(
	\bar{A}_{\psi}(s^i, a^i) - \max\limits_{a'}\bar{A}_{\psi}(s^i, a')\right)$
	\STATE $\tilde{Q^i} \gets r^i\deltat + (1 - d^i) \gamma^{\deltat} V_{\theta}(s'^i)$
	\STATE $\Delta \theta \gets \frac{1}{N}\sum\limits_{i=1}^N  \frac{\left(Q^i - \tilde{Q^i}\right)\partial_{\theta} V_{\theta}(s^i)}{\deltat}$
	\STATE $\Delta \psi \gets \frac{1}{N}\sum\limits_{i=1}^N \frac{\left(Q^i - \tilde{Q^i}\right)\partial_{\psi} \left(\bar{A}_{\psi}(s^i, a^i) - \max\limits_{a'}\bar{A}_{\psi}(s^i, a')\right) }{\deltat}$
	\STATE Update $\theta$ with \textbf{opt}$_1$, $\Delta \theta$ and learning rate $\alpha_1 \deltat$.
	\STATE Update $\psi$ with \textbf{opt}$_2$, $\Delta \psi$ and learning rate $\alpha_2 \deltat$.
	\ENDFOR
	\ENDFOR
\end{algorithmic}

>>>>

%! TEX root = icml_drau.tex
\begin{algorithmic}
	\STATE \textbf{Inputs:}
	\STATE $\theta$ and $\psi$, parameters of
	$V_{\theta}$ and $\bar{A}_{\psi}$.
	\STATE $\pi^{\text{explore}}$ and $\nu_\deltat$ defining an exploration policy.
	\STATE \textbf{opt}$_1$, \textbf{opt}$_2$, $\alpha^V \deltat$ and $\alpha^V \deltat$ optimizers and learning rates.
	\STATE $\mathcal{D}$, buffer of transitions $(s, a, r, d, s')$, with $d$ the episode termination signal.
	\STATE $\deltat$ and $\gamma$, time discretization and discount factor.
	\STATE \textbf{nb\_epochs} number of epochs.
	\STATE \textbf{nb\_steps}, number of steps per epoch.
	\STATE
	\STATE Observe initial state $s^0$
	\STATE $t \gets 0$
	\FOR {$e=0, \textbf{nb\_epochs}$}
	\FOR {$j=1, \textbf{nb\_steps}$}
	\STATE $a^k \leftarrow \pi^{\text{explore}}(s^k, \nu^k_\deltat)$.
	\STATE Perform $a^k$ and observe $(r^{k+1}, d^{k+1}, s^{k+1})$.
	\STATE Store $(s^k, a^k, r^{k+1}, d^{k+1}, s^{k+1})$ in $\mathcal{D}$.
	\STATE $k \gets k + 1$
	\ENDFOR
	\FOR {$k=0, \text{nb\_learn}$}
	\STATE \text{Sample a batch of $N$ random transitions from $\mathcal{D}$}
	\STATE $Q^i \gets V_{\theta}(s^i) + \deltat\hspace{-.17em}\left(
	\bar{A}_{\psi}(s^i, a^i) - \max\limits_{a'}\bar{A}_{\psi}(s^i, a')\right)$
	\STATE $\tilde{Q^i} \gets r^i\deltat + (1 - d^i) \gamma^{\deltat} V_{\theta}(s'^i)$
	\STATE $\Delta \theta \gets \frac{1}{N}\sum\limits_{i=1}^N  \frac{\left(Q^i - \tilde{Q^i}\right)\partial_{\theta} V_{\theta}(s^i)}{\deltat}$
	\STATE $\Delta \psi \gets \frac{1}{N}\sum\limits_{i=1}^N \frac{\left(Q^i - \tilde{Q^i}\right)\partial_{\psi} \left(\bar{A}_{\psi}(s^i, a^i) - \max\limits_{a'}\bar{A}_{\psi}(s^i, a')\right) }{\deltat}$
	\STATE Update $\theta$ with \textbf{opt}$_1$, $\Delta \theta$ and learning rate $\alpha^V \deltat$.
	\STATE Update $\psi$ with \textbf{opt}$_2$, $\Delta \psi$ and learning rate $\alpha^A \deltat$.
	\ENDFOR
	\ENDFOR
\end{algorithmic}

### 2019-01-19 paper/icml_drau/discussions.tex

\paragraph{Memory buffer size scalings}~Thm.~\ref{th:cont-params} is stated for
transitions sampled sequentially from a fixed infinite size sequence. In practice,
transitions are sampled from a memory buffer, to prevent excessive correlations.
The memory buffer used is a fixed size circular buffer, and is
filled with samples coming from a single infinite exploratory trajectory. In
our experiments, the same buffer size is used for all time discretization. When
$\deltat$ goes to $0$, this corresponds to learning nearly online on the
infinite sequence, since the buffer will only contain examples that are very
close temporally to the end of the sequence in physical time. While the resulting
algorithm \TODO{kind of} converges to a continuous time limit, it is still not
time discretization invariant, as the freshness of samples in the buffer will
vary with the time discretization.
A different choice would be to scale the size of the buffer by a factor of
$\frac{1}{\deltat}$. This would amount to always learning on the same time physical interval of
the trajectory. Our analysis, however, does not cover this case.

\paragraph{Near continuous time reinforcement learning and RMSprop}~Be careful
when using RMSprop. When you are facing a stochastic environment RMSprop
evaluates the second momentum of gradients and rescale gradients by its square
root. This second momentum will be of order δt 1 , due to the stochasticity in
the environment.  Consequently all √ gradient steps are multiplied by δt, and
learning rates need √ only be multiplied by δt instead of δt. Taking batches of
size proportional to δt 1 applies a similar fix.

>>>>

\paragraph{Memory buffer size scalings}~Thm.~\ref{th:cont-params} is stated for
transitions sampled sequentially from a fixed infinite size sequence. In practice,
transitions are sampled from a memory buffer, to prevent excessive correlations.
The memory buffer used is a fixed size circular buffer, and is
filled with samples coming from a single infinite exploratory trajectory. In
our experiments, the same buffer size is used for all time discretization. When
$\deltat$ goes to $0$, this corresponds to learning nearly online on the
infinite sequence, since the buffer will only contain examples that are very
close temporally to the end of the sequence in physical time. While the resulting
algorithm \TODO{kind of} converges to a continuous time limit, it is still not
time discretization invariant, as the freshness of samples in the buffer will
vary with the time discretization.
A different choice would be to scale the size of the buffer by a factor of
$\frac{1}{\deltat}$. This would amount to always learning on the same physical time interval of
the trajectory. Our analysis, however, does not cover this case.

\paragraph{Near continuous time reinforcement learning and RMSprop}~RMSprop~\cite{rmsprop}
estimates a batched version of the second momentum of gradients using moving
averages, and divides gradient steps by the square root of this estimate. In
deterministic environments, gradient noise typically scales as $\bigO(1)$ in
term of $\deltat$, as seen in Eq.~\eqref{eq:approx_bellman_A}.  Consequently, RMSprop
preconditioning have no effect on the order of magnitude of required learning
rates. However, in near continuous stochastic environments, as in
Eq.~\ref{eq:sde}, variance of the gradient noise typically scales as
$\bigO\left(\frac{1}{\deltat}\right)$. If batch size does not depend on
$\deltat$, RMSprop preconditioning thus multiplies gradients by a factor $\bigO(\sqrt{\deltat})$,
and learning rates need only be scaled as $\alpha \sqrt{\deltat}$ instead of
$\alpha \deltat$.

### 2019-01-19 paper/icml_drau/experiments.tex

%! TEX root = icml_drau.tex
\section{Experiments}
\label{sec:exp}
\input{lc_fig.tex}
The experiment set that we provide here is specifically aimed at showing that
the proposed method can operate efficiently on a wide range of time
discretization, without specific tuning, while standard deep Q-learning
approaches can't.  Our method, which is referred to as \emph{Deep Advantage
Updating}\TODO{Introduce the name of the algorith, earlier}, is compared to DDPG in continuous action settings and to DQN in
discrete action settings.

@@

\paragraph{Qualitative experiments: Visualizing policies and values}
To provide qualitative results, and check that robustness to time
discretization is provided both in term of raw return results, but also in term
of convergence of approximate value function and policies, we first provide results on the simple pendulum environment
provided in the openai gym classic control suite.  In the simple pendulum environment, the state space is of dimension $2$. A visualization of both the learnt value and policy functions by plotting for each point of the phase diagram $(\theta, \dot{\theta})$ its value and policy. Qualitative results are presented in
Fig.~\ref{fig:pend} for the policy and in Fig.\TODO{ref} in Supplementary \TODO{ref} for the value function.

@@

\paragraph{Quantitative experiments}
We benchmark DAU against DDPG on classic control benchmarks: Pendulum, Cartpole, BipedalWalker, Ant and Half-Cheetah environments from Open AI Gym. For Pendulum and Ant, DDPG is able to learn for the highest $\deltat$ but fails for the others, while DAU remains reasonably invariant to the time discretization. On Bipedal Walker, DDPG is never able to learn while DAU is solving the environment for every $\deltat$. On Cartpole and Half Cheetah, the results are less clear. On Cartpole, noise dominates, which makes interpretation unclear. On Half Cheetah, DAU is not clearly more invariant to time discretization than DDPG. This could be explained by the multiple suboptimal regimes that coexist in the half cheetah environment (walking on the head, walking on the back), and that DAU may fail to correctly weight the performance of each regime (as explained in the Section \ref{sec:discussions}).

>>>>

%! TEX root = icml_drau.tex
\section{Experiments}
\label{sec:exp}
\input{lc_fig.tex}
The experiment set that we provide here is specifically aimed at showing that
the proposed method DAU can operate efficiently on a wide range of time
discretization, without specific tuning, while standard deep Q-learning
approaches can't. DAU is compared to DDPG in continuous action settings and to DQN in
discrete action settings.

@@

\paragraph{Qualitative experiments: Visualizing policies and values}
To provide qualitative results, and check that robustness to time
discretization is provided both in term of raw return results, but also in term
of convergence of approximate value function and policies, we first provide results on the simple pendulum environment
from the OpenAI Gym classic control suite.  This environment's state space is of dimension $2$. A visualization of both the learnt value and policy functions is obtained by plotting for each point of the phase diagram $(\theta, \dot{\theta})$ its value and policy. results are presented in
Fig.~\ref{fig:pend} for the policy function and in Fig.\TODO{ref} in Supplementary \TODO{ref} for the value function.

@@

\paragraph{Quantitative experiments}
We benchmark DAU against DDPG on classic control benchmarks: Pendulum, Cartpole, BipedalWalker, Ant and Half-Cheetah environments from OpenAI Gym. For Pendulum and Ant, DDPG is able to learn for the highest $\deltat$ but fails for the others, while DAU remains reasonably invariant to the time discretization. On Bipedal Walker, DDPG is never able to learn while DAU is achieving decent performance for all the $\deltat$'s. On Cartpole and Half Cheetah, results are less clear. On Cartpole, noise dominates, which makes interpretation difficult. On Half Cheetah, DAU is not clearly more invariant to time discretization than DDPG. This could be explained by the multiple suboptimal regimes that coexist in the half cheetah environment (walking on the head, walking on the back), and that DAU may fail to correctly weight the performance of each regime (as explained in Section \ref{sec:discussions}).

### 2019-01-19 paper/icml_drau/framework.tex

Many reinforcement learning environments are not intrinsically
time-discrete, but discretizations of an underlying continuous-time
environment. For instance, many simulated control environments, such as
the Mujoco environments~\cite{ddpg} or OpenAI Gym classic control
environments~\cite{gym}, are discretization of continuous-time control
problems.  In simulated environments, the time discretization is fixed by
the simulator, and is often used to approximate an underlying
differential equation.  \TODO{(For example) $->$ In this case}, the timestep may correspond to the
number of frames generated by second.  In real world environments,
sensors and actuators have a fixed time precision: cameras can only
capture a fixed amount of frames per second, and physical limitations
prevent actuators from responding instantaneously. The quality of these
components thus imposes a lower bound on the discretization timestep. As
the timestep $\deltat$ is not an intrinsic parameter of the
environment, but a constraint imposed by computational ressources, we
would expect that decreasing $\deltat$ could only improve the performance
of RL agents \TODO{We would remove: (though it might make optimization
harder) It is not in the spririt of what is said here}.  RL algorithms should, at least, be resilient to a change of
$\deltat$, and should remain viable when $\deltat \rightarrow 0$.
Besides, designing a time discretization invariant algorithm could
alleviate tedious hyperparameterization by providing better defaults for
time-horizon-related parameters.

>>>>

Many reinforcement learning environments are not intrinsically
time-discrete, but discretizations of an underlying continuous-time
environment. For instance, many simulated control environments, such as
the Mujoco environments~\cite{ddpg} or OpenAI Gym classic control
environments~\cite{gym}, are discretization of continuous-time control
problems.  In simulated environments, the time discretization is fixed by
the simulator, and is often used to approximate an underlying
differential equation.  \TODO{(For example) $->$ In this case}, the timestep may correspond to the
number of frames generated by second.  In real world environments,
sensors and actuators have a fixed time precision: cameras can only
capture a fixed amount of frames per second, and physical limitations
prevent actuators from responding instantaneously. The quality of these
components thus imposes a lower bound on the discretization timestep. As
the timestep $\deltat$ is not an intrinsic parameter of the
environment, but a constraint imposed by computational ressources, we
would expect that decreasing $\deltat$ could only improve the performance
of RL agents \TODO{We would remove: (though it might make optimization
harder) It is not in the spririt of what is said here. NDY keep it.
Anyone with a bit of experience in traditional optim would scream and say exactly
that, if it was not included.}.  RL algorithms should, at least, be resilient to a change of
$\deltat$, and should remain viable when $\deltat \rightarrow 0$.
Besides, designing a time discretization invariant algorithm could
alleviate tedious hyperparameterization by providing better defaults for
time-horizon-related parameters.
